Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-04-03 WRITEUP/ms.tex

% ------------------
\section{Numerical simulations}\label{sec:sim}
The simulations used in this work are described by \citealt{Pallottini17a, Pallottini17b} and are only briefly summarized here.
%
\ncode{Serra}\footnote{Greenhouse in Italian} is a suite of cosmological zoom-in simulations performed using Eulerian hydrodynamics and adaptive mesh refinement (AMR) techniques to achieve high spatial resolution in regions of interest (e.g., regions of high density).
%
In particular, it uses a modified version of \ncode{ramses}
\citep{Teyssier02a} as the AMR backend. The simulation used here
covers a comoving box of 20\,Mpc $h$\pmOne in size. The simulation
zooms in on a target halo of mass $M_{\rm DM}\simeq10^{11} \Msun$. The Lagrangian
region of the halo (2.1\,Mpc $h$\pmOne) has a dark matter
%mm (equivalent baryonic) 
mass resolution of $\simeq 6\times 10^4 \Msun$, 
%mm ($1.2 \times 10^4 \Msun$) [expanded out the meaning, but check if
%I have it right]  Such
     equivalent at initial density to a baryonic mass resolution of $1.2
     \times 10^4 \Msun$.  This 
region is spatially refined with a quasi-Lagrangian criterion based on
a mass threshold, so that a cell is refined if its total dark matter
plus baryonic mass exceeds the mass resolution by a factor of
eight. The final finest refined cell in the zoom in region has size
$l_{\rm cell}\simeq$\,30\,pc (at $z = 6$). 
%mm
%    \MM{Is this really the resolution of dark matter particles?  The baryonic mass (i.e.\ gas) resolution depends on the 
%    local density, which would have to be specified here.} \AP{the setting description has been extended}
     Cell sizes are comparable to the sizes of local giant molecular clouds \citep[e.g.,][]{Sanders85a, Federrath13a, Goodman14a},
%mm 
%     \MM{Nota Bene!} \AP{This note is discussed in details in the following section. In the simulation we do not treat 5--10 cells as MCs, so I would remove the point from the present} although the resolved scale of 5--10 cells is rather larger.
The models include a chemical network 
following e$^{-}$, H$^+$, H$^-$, He, He$^+$, He$^{++}$, H$_2$, and
H$_2^+$ \citep{Grassi14a,Bovino16a}. Of particular importance here is
that abundances are calculated using an on-the-fly non-equilibrium
formation of molecular hydrogen scheme  
%mm following \citet{Pallotini17b} [the order had to be reversed for alphabetization]
    described by \citet{Pallottini17a}.

@@

One of the main advantages for studying the dynamical properties of simulated galaies is that we can examine how their properties evolve with time.
%
%This is advantageous especially at early cosmic epochs, when the densest structures are beginning to form; gas is constantly being accreted onto the central galaxy from the cosmic web and satellite galaxies, thereby leading to bursts of \SF. Meanwhile, tidal forces resulting from interactions with these surrounding galaxies can disrupt the main disk and arms, likely leading to different dynamical states for the molecular structures compared to more evolved galaxies found at a later cosmic time (e.g., some molecular structures may disperse while others may agglomerate into more massive ones). % mainly expecting differences in alpha_vir, sigma, M_cl.
%mm [delete para break]
The \SF history\footnote{Note that here the SFR is calculated based on
  the stellar mass formed in the last 10 Myr within 3.5 kpc from the
  galaxy center of mass. The SFR plotted in Figure 2 of
  \citet{Pallottini17b} is a factor of two higher since there the SFR
  accounts for the contribution from massive satellite galaxies within
  the virial radius ($\approx$15\,kpc).} of \flower is shown in
\Fig{SFH}. The SFR of \flower varies between $\sim$30--80 \Msun\,yr\pmOne as it evolves from an actively accreting phase to a starburst phase after a merger, and then back to a relatively quiescent phase, over the simulated $\approx 700$ Myr.

@@

The highest density threshold of $n_{\rm cut}$\eq31.62\,\cc corresponds to 
a minimum MCC mass of the order of 10$^{5.5}$\,\Msun for the densest structure.
However, not all evolutionary stages considered have at least one MCC at this highest density threshold, 
we therefore consider MCCs identified with the second highest density threshold  ($n_{\rm cut}$\eq18.96\,\cc) in the following. 
We show in \Fig{dist} the MCC distribution in terms of their molecular gas mass, radius, and gas mass fraction, which we define as
\begin{equation}
f_{\rm gas} = \frac{M_{\rm cl}} {\left(M_{\rm cl} + M_\star\right)},
\end{equation}
where $M_\star$ is the stellar mass within the MCC volume.
%
We note that the distributions vary for different evolutionary stages of \flower. 
%
The massive ($M_{\rm cl}>$10$^8$\,\Msun) 
clouds identified at low density cut (see top left panel of \Fig{dist})  
correspond to the molecular disk of the
galaxy, which is identified as a single component for low $n_{\rm
  cut}$. The mass of the MCCs identified ranges between $M_{\rm
  cl}\simeq$10$^{5.5-8.5}$\,\Msun, consistent with those observed in
\z$\sim$2 galaxies in rest-frame UV and optical light
\citep{Elmegreen07a, Elmegreen09a}. \AF{you do not comment on $R$ and
  $f_{gas}$ here..Also, what is the total number of MCCs?} \MM{In the
  high-density plot, it looks like you have quite a number of
  unresolved clouds with $R < 40$~pc. Any idea what's going on there?}

@@

MCCs in \flower\ are characterized by large velocity dispersions ($\sigma \simeq 100 {\rm km s}^{-1}$) and sizes ($R\simeq 100$ pc). These values are consistent with those found in starburst galaxies as SMM J2135-0102, located at $z=...$ \AF{Give refs. here and in caption; in addition it would good to give the properties of this galaxy}. 
Sizes are somewhat dependent on the choice of $n_{\rm cut}$. As such density threshold is increased, some of the MCCs within the main disk break into multiple sub-MCCs. In this case, we effectively identify a population of denser molecular structures (see \Fig{MCC}). Instead, we find that $\sigma$ is rather insensitive to the actual value of $n_{\rm cut}$.

@@

% Starburst 
Some MCCs in the starburst phase have lower velocity dispersion, $\alpha_{\rm vir}$, and $\sigma^2/R$ (closer to loci with lower external pressure) compared to the accreting phase. \DL{Is this significant enough to say something physical? If so, not immediately clear to me what cause the differences.}
% PVE
We also compare the MCCs identified in our simulations to those observed in the Milky Way in the context of the $\sigma^2/R - \Sigma$ relation (see bottom panels of \Fig{larsons_single}). Dashed lines in the figure show the loci along which the given external pressures are needed for MCCs to have certain linewidths for a given set of surface densities (see \Sec{PVE}).
% (Rice+16) find variation in $V_0^2$, where they find larger linewidths for GMCCs in inner galactic disk, but they find that the clouds in the Galactic disk are still consistent with SVE (see also Solomon87a ; cf. Heyer+09).
The MCCs in the satellite galaxies are found to follow the locus of
$\log{(P/\textrm{K cm}^{-3})}$\eq6, \MM{This is not shown in the
  figure.  If you want to claim it, it needs to be shown somewhere.} which
is consistent with that found for the bulk of dense gas in the simulation in the phase-space analysis presented in \citet{Pallottini17a}.

@@

We compare the dynamics of MCCs in \flower to \obs in the context of
the Larson's relation. \MM{Which one?}
The MCCs of \flower are found to have higher $\sigma$ and $\Sigma$
\MM{compared to Milky Way clouds?}
systematically regardless of the 
density threshold $n_{\rm cut}$ adopted in identifying the structures.
The scaling relations found are insensitive to the various density
cuts.  \MM{This point needs to be made more clearly, probably by
  defining the colors that correspond to the density cuts in the figures.}
 The velocity dispersion remains $\gtrsim$100\,\kms even when we increase the density threshold and 
 even for the molecular substructures.  \MM{You never clearly identify
   what you mean by ``molecular substructures.''  Do you just mean
   objects found with high $n_{\rm cut}$?}
 Such a high velocity dispersion likely results from the strong supernova and stellar feedback \flower experienced 
 over the multiple episodes of bursty \SF.   \MM{Or is it from the
   deep potential?  Feedback-driven turbulence famously decays quickly.}

@@

Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
Observations leveraging the combination of spatial-spectral imaging of
multi-band continuum and spectral line emission are crucial for better understanding
the role of \highz galaxy populations
in the context of galaxy evolution and the ISM physics behind their intense star formation in the early Universe.
Cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics, and subject to the sub-grid models adopted,
serve as a useful tool for examining and making predictions on the
morphology and dynamics of 
the molecular ISM of the first galaxies.
%Future higher resolution zoom-in simulations would certainly be useful to understand
%the evolving dynamical properties of the lower-mass molecular structures in
%galaxies at EoR under the influence of gas inflow/outflow (i.e., accounting for energy injection at large scale due to infall/outflow).
These predictions can in turn be
tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
enabling us to test our findings and refine future simulations to shed light on the physical processes behind
\SF since the cosmic dark ages.

>>>>

% ------------------
\section{Numerical simulations}\label{sec:sim}
The simulations used in this work are described by \citealt{Pallottini17a, Pallottini17b} and are only briefly summarized here.
%
\ncode{Serra}\footnote{Greenhouse in Italian.} is a suite of cosmological zoom-in simulations performed using Eulerian hydrodynamics and adaptive mesh refinement (AMR) techniques to achieve high spatial resolution in regions of interest (e.g., regions of high density).
%
In particular, it uses a modified version of \ncode{ramses}
\citep{Teyssier02a} as the AMR backend. The simulation used here
covers a comoving box of 20\,Mpc $h$\pmOne in size. The simulation
zooms in on a target halo of mass $M_{\rm DM}\simeq10^{11} \Msun$. The Lagrangian
region of the halo (2.1\,Mpc $h$\pmOne) has a dark matter
%mm (equivalent baryonic) 
mass resolution of $\simeq 6\times 10^4 \Msun$, 
%mm ($1.2 \times 10^4 \Msun$) [expanded out the meaning, but check if
%I have it right]  Such
     equivalent at initial density to a baryonic mass resolution of $1.2
     \times 10^4 \Msun$.  This 
region is spatially refined with a quasi-Lagrangian criterion based on
a mass threshold, so that a cell is refined if its total dark matter
plus baryonic mass exceeds the mass resolution by a factor of
eight. The final finest refined cell in the zoom in region has size
$l_{\rm cell}\simeq$\,30\,pc (at $z = 6$). 
%mm
%    \MM{Is this really the resolution of dark matter particles?  The baryonic mass (i.e.\ gas) resolution depends on the 
%    local density, which would have to be specified here.} \AP{the setting description has been extended}
     Cell sizes are comparable to the sizes of local giant molecular clouds \citep[e.g.,][]{Sanders85a, Federrath13a, Goodman14a},
%mm 
%     \MM{Nota Bene!} \AP{This note is discussed in details in the following section. In the simulation we do not treat 5--10 cells as MCs, so I would remove the point from the present} although the resolved scale of 5--10 cells is rather larger.
The models include a chemical network 
following e$^{-}$, H$^+$, H$^-$, He, He$^+$, He$^{++}$, H$_2$, and
H$_2^+$ \citep{Grassi14a,Bovino16a}. Of particular importance here is
that abundances are calculated using an on-the-fly non-equilibrium
formation of molecular hydrogen scheme  
%mm following \citet{Pallotini17b} [the order had to be reversed for alphabetization]
    described by \citet{Pallottini17a}.

@@

One of the main advantages for studying the dynamical properties of simulated galaies is that we can examine how their properties evolve with time.
%
%This is advantageous especially at early cosmic epochs, when the densest structures are beginning to form; gas is constantly being accreted onto the central galaxy from the cosmic web and satellite galaxies, thereby leading to bursts of \SF. Meanwhile, tidal forces resulting from interactions with these surrounding galaxies can disrupt the main disk and arms, likely leading to different dynamical states for the molecular structures compared to more evolved galaxies found at a later cosmic time (e.g., some molecular structures may disperse while others may agglomerate into more massive ones). % mainly expecting differences in alpha_vir, sigma, M_cl.
%mm [delete para break]
The \SF history\footnote{Note that here the SFR is calculated based on
  the stellar mass formed in the last 4 Myr within 3.5 kpc from the
  galaxy center of mass. The SFR plotted in Figure 2 of
  \citet{Pallottini17b} is a factor of two higher since there the SFR
  accounts for the contribution from massive satellite galaxies within
  the virial radius ($\approx$15\,kpc).} of \flower is shown in
\Fig{SFH}. The SFR of \flower varies between $\sim$30--80 \Msun\,yr\pmOne as it evolves from an actively accreting phase to a starburst phase after a merger, and then back to a relatively quiescent phase, over the simulated $\approx 700$ Myr.

@@

The highest density threshold of $n_{\rm cut}$\eq31.62\,\cc corresponds to 
a minimum MCC mass of the order of 10$^{5.5}$\,\Msun for the densest structure.
However, not all evolutionary stages considered have at least one MCC at this highest density threshold, 
we therefore consider MCCs identified with the second highest density threshold  ($n_{\rm cut}$\eq18.96\,\cc) in the following. 
We show in \Fig{dist} the MCC distribution in terms of their molecular gas mass, radius, and gas mass fraction, which we define as
\begin{equation}
f_{\rm gas} = \frac{M_{\rm cl}} {\left(M_{\rm cl} + M_\star\right)},
\end{equation}
where $M_\star$ is the stellar mass within the MCC volume.
%
We note that the distributions vary for different evolutionary stages of \flower. 
%
The massive ($M_{\rm cl}>$10$^8$\,\Msun) 
clouds identified at low density cut (see top left panel of \Fig{dist})  
correspond to the molecular disk of the
galaxy, which is identified as a single component for low $n_{\rm
  cut}$. The mass of the MCCs identified ranges between $M_{\rm
  cl}\simeq$10$^{5.5-8.5}$\,\Msun, consistent with those observed in
\z$\sim$2 galaxies in rest-frame UV and optical light
\citep{Elmegreen07a, Elmegreen09a}. \AF{you do not comment on $R$ and
  $f_{gas}$ here..Also, what is the total number of MCCs?}

@@

MCCs in \flower\ are characterized by large velocity dispersions ($\sigma \simeq 100 {\rm km s}^{-1}$) and sizes ($R\simeq 100$ pc). These values are consistent with those found in starburst galaxies as SMM J2135-0102, located at $z=...$ \AF{Give refs. here and in caption; in addition it would good to give the properties of this galaxy}. 
Sizes are somewhat dependent on the choice of $n_{\rm cut}$. As such density threshold is increased, some of the MCCs within the main disk break into multiple sub-MCCs. In this case, we effectively identify a population of denser molecular structures (see \Fig{MCC}). On the other hand, we find that $\sigma$ is rather insensitive to the actual value of $n_{\rm cut}$.

@@

% Starburst 
Some MCCs in the starburst phase have lower velocity dispersion, $\alpha_{\rm vir}$, and $\sigma^2/R$ compared to the accreting phase. \DL{Is this significant enough to say something physical? If so, not immediately clear to me what cause the differences.}
% PVE
We also compare the MCCs identified in our simulations to those observed in the Milky Way in the context of the $\sigma^2/R - \Sigma$ relation (see bottom panels of \Fig{larsons_single}). Dashed lines in the figure show the loci along which the given external pressures are needed for MCCs to have certain linewidths for a given set of surface densities (see \Sec{PVE}).
% (Rice+16) find variation in $V_0^2$, where they find larger linewidths for GMCCs in inner galactic disk, but they find that the clouds in the Galactic disk are still consistent with SVE (see also Solomon87a ; cf. Heyer+09).
The MCCs in the satellite galaxies are found to follow the locus of
$\log{(P/\textrm{K cm}^{-3})}$\eq6, \MM{This is not shown in the
  figure.  If you want to claim it, it needs to be shown somewhere.} which
is consistent with that found for the bulk of dense gas in the simulation in the phase-space analysis presented in \citet{Pallottini17a}.

@@

We compare the dynamics of MCCs in \flower to \obs in the context of
the Larson's relation. \MM{Which one?}
The MCCs of \flower are found to have higher $\sigma$ and $\Sigma$
\MM{compared to Milky Way clouds?}
systematically regardless of the 
density threshold $n_{\rm cut}$ adopted in identifying the structures.
The scaling relations found are insensitive to the various density
cuts.  \MM{This point needs to be made more clearly, probably by
  defining the colors that correspond to the density cuts in the figures.} \DL{They are color-coded by ncut in Fig 5 and 7.}
 The velocity dispersion remains $\gtrsim$100\,\kms even when we increase the density threshold and 
 even for the molecular substructures.  \MM{You never clearly identify
   what you mean by ``molecular substructures.''  Do you just mean
   objects found with high $n_{\rm cut}$?}
 Such a high velocity dispersion likely results from the strong supernova and stellar feedback \flower experienced 
 over the multiple episodes of bursty \SF.   \MM{Or is it from the
   deep potential?  Feedback-driven turbulence famously decays quickly.}

@@

Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
Observations leveraging the combination of spatial-spectral imaging of
multi-band continuum and spectral line emission are crucial for better understanding
the role of \highz galaxy populations
in the context of galaxy evolution and the ISM physics behind their intense star formation in the early Universe.
Cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics and are subject to the sub-grid models adopted,
serve as a useful tool for examining and making predictions on the
morphology and dynamics of 
the molecular ISM of the first galaxies.
%Future higher resolution zoom-in simulations would certainly be useful to understand
%the evolving dynamical properties of the lower-mass molecular structures in
%galaxies at EoR under the influence of gas inflow/outflow (i.e., accounting for energy injection at large scale due to infall/outflow).
These predictions can in turn be
tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
enabling us to test our findings and refine future simulations to shed light on the physical processes behind
\SF since the cosmic dark ages.
